Equations & R Code

### 2017-03-22 manuscript/coexistence-stability_SI.Rmd
updates all analyses with truncation and low atol for no errors or warnings

  require('deSolve') # for solving continuous differential equations
  require('mvtnorm') # for multivariate normal distribution functions

  ##  Assign parameter values to appropriate lists
  DNR <- c(D=c(D1,D2,D3,D4),   # initial dormant state abundance
           N=c(N1,N2,N3,N4),   # initial live state abundance
           R=R)                # initial resource level

@@

  ##  Resource uptake function (Hill function)
  uptake_R <- function(r, R, a, b) {
    return((r*R^a) / (b^a + R^a))
  }

@@

  ####
  #### Simulate model -----------------------------------------------------
  ####
  days           <- c(1:days_to_track)
  num_spp        <- length(parms$r)
  nmsDNR         <- names(DNR)
  dormants       <- grep("D", names(DNR))
  NR             <- DNR[-dormants] 
  nmsNR          <- names(NR)
  gVec           <- getG(sigE = sigE, rho = rho, nTime = seasons, num_spp = num_spp)
  Rvector        <- rlnorm(seasons, Rmu, Rsd_annual)
  saved_outs     <- matrix(ncol=length(DNR), nrow=seasons+1)
  saved_outs[1,] <- DNR

  ##  Loop over seasons
  for(season_now in 1:seasons) {
    # Simulate continuous growing  season
    output   <- ode(y = NR, times=days, func = updateNR, parms = parms)
    NR       <- output[nrow(output),nmsNR]
    dormants <- grep("D", names(DNR)) 
    DNR      <- c(DNR[dormants], NR)

>>>>

  require('deSolve') # for solving continuous differential equations
  require('mvtnorm') # for multivariate normal distribution functions
  require('Runuran') # for truncated log normal random number generator

  ##  Assign parameter values to appropriate lists
  DNR <- c(D=c(D1,D2,D3,D4),   # initial dormant state abundance
           N=c(N1,N2,N3,N4),   # initial live state abundance
           R=R)                # initial resource level
  DNR_inits <- DNR

@@

  ##  Resource uptake function (Hill function)
  uptake_R <- function(r, R, a, b) {
    return( (r*R^a) / (b^a + R^a) )
  }

@@

  ####
  #### Simulate model -----------------------------------------------------
  ####
  days           <- c(1:days_to_track)
  num_spp        <- length(parms$r)
  nmsDNR         <- names(DNR)
  dormants       <- grep("D", names(DNR))
  NR             <- DNR[-dormants] 
  nmsNR          <- names(NR)
  gVec           <- getG(sigE = sigE, rho = rho, nTime = seasons, num_spp = num_spp)
  Rvector        <- urlnorm(seasons, Rmu, Rsd_annual, lb = 0, ub = 200)
  saved_outs     <- matrix(ncol=length(DNR), nrow=seasons+1)
  saved_outs[1,] <- DNR

  ##  Loop over seasons
  for(season_now in 1:seasons) {
    # Simulate continuous growing  season
    output   <- ode(y = NR, times = days, func = updateNR, parms = parms, atol = 1e-100)
    NR       <- output[nrow(output),nmsNR]
    dormants <- grep("D", names(DNR)) 
    DNR      <- c(DNR[dormants], NR)

### 2017-03-22 manuscript/coexistence-stability_SI.tex
updates all analyses with truncation and low atol for no errors or warnings

  \NormalTok{##  Assign parameter values to appropriate lists}
  \NormalTok{DNR <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DataTypeTok{D=}\KeywordTok{c}\NormalTok{(D1,D2,D3,D4),   }\CommentTok{# initial dormant state abundance}
           \DataTypeTok{N=}\KeywordTok{c}\NormalTok{(N1,N2,N3,N4),   }\CommentTok{# initial live state abundance}
           \DataTypeTok{R=}\NormalTok{R)                }\CommentTok{# initial resource level}

@@

  \NormalTok{##  Generate germination fractions}
  \NormalTok{getG <-}\StringTok{ }\NormalTok{function(sigE, rho, nTime, num_spp) \{}
    \NormalTok{varcov       <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\KeywordTok{rep}\NormalTok{(rho*sigE,num_spp*}\DecValTok{2}\NormalTok{), num_spp, num_spp)}
    \KeywordTok{diag}\NormalTok{(varcov) <-}\StringTok{ }\NormalTok{sigE}
    \NormalTok{if(sigE >}\StringTok{ }\DecValTok{0}\NormalTok{) \{ varcov <-}\StringTok{ }\NormalTok{Matrix::}\KeywordTok{nearPD}\NormalTok{(varcov)$mat \} }\CommentTok{# crank through nearPD... }
                                                          \CommentTok{# ...to fix rounding errors }
    \NormalTok{varcov <-}\StringTok{ }\KeywordTok{as.matrix}\NormalTok{(varcov)}
    \NormalTok{e      <-}\StringTok{ }\KeywordTok{rmvnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{nTime, }\DataTypeTok{mean =} \KeywordTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{,num_spp), }\DataTypeTok{sigma =} \NormalTok{varcov)}
    \NormalTok{g      <-}\StringTok{ }\KeywordTok{exp}\NormalTok{(e) /}\StringTok{ }\NormalTok{(}\DecValTok{1}\NormalTok{+}\KeywordTok{exp}\NormalTok{(e))}
    \KeywordTok{return}\NormalTok{(g)}
  \NormalTok{\}}

  \NormalTok{####}
  \NormalTok{#### Simulate model -----------------------------------------------------}
  \NormalTok{####}
  \NormalTok{days           <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{:days_to_track)}
  \NormalTok{num_spp        <-}\StringTok{ }\KeywordTok{length}\NormalTok{(parms$r)}
  \NormalTok{nmsDNR         <-}\StringTok{ }\KeywordTok{names}\NormalTok{(DNR)}
  \NormalTok{dormants       <-}\StringTok{ }\KeywordTok{grep}\NormalTok{(}\StringTok{"D"}\NormalTok{, }\KeywordTok{names}\NormalTok{(DNR))}
  \NormalTok{NR             <-}\StringTok{ }\NormalTok{DNR[-dormants] }
  \NormalTok{nmsNR          <-}\StringTok{ }\KeywordTok{names}\NormalTok{(NR)}
  \NormalTok{gVec           <-}\StringTok{ }\KeywordTok{getG}\NormalTok{(}\DataTypeTok{sigE =} \NormalTok{sigE, }\DataTypeTok{rho =} \NormalTok{rho, }\DataTypeTok{nTime =} \NormalTok{seasons, }\DataTypeTok{num_spp =} \NormalTok{num_spp)}
  \NormalTok{Rvector        <-}\StringTok{ }\KeywordTok{urlnorm}\NormalTok{(seasons, Rmu, Rsd_annual, }\DataTypeTok{lb =} \DecValTok{0}\NormalTok{, }\DataTypeTok{ub =} \DecValTok{200}\NormalTok{)}
  \NormalTok{saved_outs     <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{ncol =} \KeywordTok{length}\NormalTok{(DNR), }\DataTypeTok{nrow =} \NormalTok{seasons}\DecValTok{+1}\NormalTok{)}
  \NormalTok{saved_outs[}\DecValTok{1}\NormalTok{,] <-}\StringTok{ }\NormalTok{DNR }

  \NormalTok{##  Loop over seasons}
  \NormalTok{for(season_now in }\DecValTok{1}\NormalTok{:seasons) \{}
    \CommentTok{# Simulate continuous growing  season}
    \NormalTok{output   <-}\StringTok{ }\KeywordTok{ode}\NormalTok{(}\DataTypeTok{y =} \NormalTok{NR, }\DataTypeTok{times =} \NormalTok{days, }\DataTypeTok{func =} \NormalTok{updateNR, }\DataTypeTok{parms =} \NormalTok{parms)}
    \NormalTok{NR       <-}\StringTok{ }\NormalTok{output[}\KeywordTok{nrow}\NormalTok{(output),nmsNR]}
    \NormalTok{dormants <-}\StringTok{ }\KeywordTok{grep}\NormalTok{(}\StringTok{"D"}\NormalTok{, }\KeywordTok{names}\NormalTok{(DNR)) }
    \NormalTok{DNR      <-}\StringTok{ }\KeywordTok{c}\NormalTok{(DNR[dormants], NR)}

>>>>

  \NormalTok{##  Assign parameter values to appropriate lists}
  \NormalTok{DNR <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DataTypeTok{D=}\KeywordTok{c}\NormalTok{(D1,D2,D3,D4),   }\CommentTok{# initial dormant state abundance}
           \DataTypeTok{N=}\KeywordTok{c}\NormalTok{(N1,N2,N3,N4),   }\CommentTok{# initial live state abundance}
           \DataTypeTok{R=}\NormalTok{R)                }\CommentTok{# initial resource level}
  \NormalTok{DNR_inits <-}\StringTok{ }\NormalTok{DNR}

@@

  \NormalTok{##  Generate germination fractions}
  \NormalTok{getG <-}\StringTok{ }\NormalTok{function(sigE, rho, nTime, num_spp) \{}
    \NormalTok{varcov       <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\KeywordTok{rep}\NormalTok{(rho*sigE,num_spp*}\DecValTok{2}\NormalTok{), num_spp, num_spp)}
    \KeywordTok{diag}\NormalTok{(varcov) <-}\StringTok{ }\NormalTok{sigE}
    \NormalTok{if(sigE >}\StringTok{ }\DecValTok{0}\NormalTok{) \{ varcov <-}\StringTok{ }\NormalTok{Matrix::}\KeywordTok{nearPD}\NormalTok{(varcov)$mat \} }\CommentTok{# crank through nearPD to fix rounding errors }
    \NormalTok{varcov <-}\StringTok{ }\KeywordTok{as.matrix}\NormalTok{(varcov)}
    \NormalTok{e      <-}\StringTok{ }\KeywordTok{rmvnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{nTime, }\DataTypeTok{mean =} \KeywordTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{,num_spp), }\DataTypeTok{sigma =} \NormalTok{varcov)}
    \NormalTok{g      <-}\StringTok{ }\KeywordTok{exp}\NormalTok{(e) /}\StringTok{ }\NormalTok{(}\DecValTok{1}\NormalTok{+}\KeywordTok{exp}\NormalTok{(e))}
    \KeywordTok{return}\NormalTok{(g)}
  \NormalTok{\}}

  \NormalTok{####}
  \NormalTok{#### Simulate model -----------------------------------------------------}
  \NormalTok{####}
  \NormalTok{days           <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{:days_to_track)}
  \NormalTok{num_spp        <-}\StringTok{ }\KeywordTok{length}\NormalTok{(parms$r)}
  \NormalTok{nmsDNR         <-}\StringTok{ }\KeywordTok{names}\NormalTok{(DNR)}
  \NormalTok{dormants       <-}\StringTok{ }\KeywordTok{grep}\NormalTok{(}\StringTok{"D"}\NormalTok{, }\KeywordTok{names}\NormalTok{(DNR))}
  \NormalTok{NR             <-}\StringTok{ }\NormalTok{DNR[-dormants] }
  \NormalTok{nmsNR          <-}\StringTok{ }\KeywordTok{names}\NormalTok{(NR)}
  \NormalTok{gVec           <-}\StringTok{ }\KeywordTok{getG}\NormalTok{(}\DataTypeTok{sigE =} \NormalTok{sigE, }\DataTypeTok{rho =} \NormalTok{rho, }\DataTypeTok{nTime =} \NormalTok{seasons, }\DataTypeTok{num_spp =} \NormalTok{num_spp)}
  \NormalTok{Rvector        <-}\StringTok{ }\KeywordTok{urlnorm}\NormalTok{(seasons, Rmu, Rsd_annual, }\DataTypeTok{lb =} \DecValTok{0}\NormalTok{, }\DataTypeTok{ub =} \DecValTok{100}\NormalTok{)}
  \NormalTok{saved_outs     <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{ncol=}\KeywordTok{length}\NormalTok{(DNR), }\DataTypeTok{nrow=}\NormalTok{seasons}\DecValTok{+1}\NormalTok{)}
  \NormalTok{saved_outs[}\DecValTok{1}\NormalTok{,] <-}\StringTok{ }\NormalTok{DNR }

  \NormalTok{##  Loop over seasons}
  \NormalTok{for(season_now in }\DecValTok{1}\NormalTok{:seasons) \{}
    \CommentTok{# Simulate continuous growing  season}
    \NormalTok{output   <-}\StringTok{ }\KeywordTok{ode}\NormalTok{(}\DataTypeTok{y =} \NormalTok{NR, }\DataTypeTok{times =} \NormalTok{days, }\DataTypeTok{func =} \NormalTok{updateNR, }\DataTypeTok{parms =} \NormalTok{parms, }\DataTypeTok{atol =} \FloatTok{1e-100}\NormalTok{)}
    \NormalTok{NR       <-}\StringTok{ }\NormalTok{output[}\KeywordTok{nrow}\NormalTok{(output),nmsNR]}
    \NormalTok{dormants <-}\StringTok{ }\KeywordTok{grep}\NormalTok{(}\StringTok{"D"}\NormalTok{, }\KeywordTok{names}\NormalTok{(DNR)) }
    \NormalTok{DNR      <-}\StringTok{ }\KeywordTok{c}\NormalTok{(DNR[dormants], NR)}

### 2017-03-22 manuscript/coexistence-stability_ms.Rmd
updates all analyses with truncation and low atol for no errors or warnings

We assume the resource pool is not replenished within a growing season.
Resource replenishment occurs between growing seasons, and the resource pool (*R*) at the start of the growing season is $R_{\tau(0)} = R^+$, where $R^+$ is a random resource pulse drawn from a log-normal distribution with mean $\mu(R^+)$ and standard deviation $\sigma(R^+)$. 
\new{Taken all together, we can combine equations 1 and 4 to define the discrete transitions between live and dormant biomass at the end of a growing season.
Thus, the initial conditions for each state ($D, N, R$) at the beginning of growing season $\tau+1$ are:}
\vspace{-3em}
\begin{align}
  D_{i,\tau(0) + 1} &= (1-\gamma_{i,\tau}) \left[\alpha_i N_{i,\tau(T)} + D_{i,\tau(T)} \right] (1-\eta_i) \\
  N_{i,\tau(0) + 1} &= \gamma_{i,\tau} \left[\alpha_i N_{i,\tau(T)} + D_{i,\tau(T)} \right] (1-\eta_i) \\
  R_{\tau + 1} &= \text{lognormal}\left(\mu(R^+), \sigma(R^+) \right),
\end{align}
\vspace{-3em}
\noindent \new{where, as above, $\tau(T)$ denotes the end of growing season $\tau$ and $\tau(0) + 1$ denotes the beginning of growing season $\tau+1$.} 
Model parameters and notation are described in Table 1.

>>>>

We assume the resource pool is not replenished within a growing season.
Resource replenishment occurs between growing seasons, and the resource pool (*R*) at the start of the growing season is $R_{\tau(0)} = R^+$, where $R^+$ is a random resource pulse drawn from a log-normal distribution with mean $\mu(R^+)$ and standard deviation $\sigma(R^+)$. 
\new{Taken all together, we can combine equations 1 and 4 to define the discrete transitions between live and dormant biomass at the end of a growing season.
Thus, the initial conditions for each state ($D, N, R$) at the beginning of growing season $\tau+1$ are:}
\vspace{-3em}
\begin{align}
  D_{i,\tau(0) + 1} &= (1-\gamma_{i,\tau}) \left[\alpha_i N_{i,\tau(T)} + D_{i,\tau(T)} \right] (1-\eta_i) \\
  N_{i,\tau(0) + 1} &= \gamma_{i,\tau} \left[\alpha_i N_{i,\tau(T)} + D_{i,\tau(T)} \right] (1-\eta_i) \\
  R_{\tau + 1} &= \text{lognormal}\left(\mu(R^+), \sigma(R^+) \right)_0^{200},
\end{align}
\vspace{-3em}
\noindent \new{where, as above, $\tau(T)$ denotes the end of growing season $\tau$ and $\tau(0) + 1$ denotes the beginning of growing season $\tau+1$.} 
\new{The subscript (0) and superscript (200) indicates a lognormal distribution truncated at those values.
We used the function \texttt{urlnorm} from the Runuran package (CITATION) to generate values from the truncated lognormal distribution.}
Model parameters and notation are described in Table 1.

### 2017-03-22 manuscript/coexistence-stability_ms.tex
updates all analyses with truncation and low atol for no errors or warnings

\begin{align}
  D_{i,\tau(0) + 1} &= (1-\gamma_{i,\tau}) \left[\alpha_i N_{i,\tau(T)} + D_{i,\tau(T)} \right] (1-\eta_i) \\
  N_{i,\tau(0) + 1} &= \gamma_{i,\tau} \left[\alpha_i N_{i,\tau(T)} + D_{i,\tau(T)} \right] (1-\eta_i) \\
  R_{\tau + 1} &= \text{lognormal}\left(\mu(R^+), \sigma(R^+) \right),
\end{align}\vspace{-3em}

\noindent \new{where, as above, $\tau(T)$ denotes the end of growing season $\tau$ and $\tau(0) + 1$ denotes the beginning of growing season $\tau+1$.}
Model parameters and notation are described in Table 1.

>>>>

\begin{align}
  D_{i,\tau(0) + 1} &= (1-\gamma_{i,\tau}) \left[\alpha_i N_{i,\tau(T)} + D_{i,\tau(T)} \right] (1-\eta_i) \\
  N_{i,\tau(0) + 1} &= \gamma_{i,\tau} \left[\alpha_i N_{i,\tau(T)} + D_{i,\tau(T)} \right] (1-\eta_i) \\
  R_{\tau + 1} &= \text{lognormal}\left(\mu(R^+), \sigma(R^+) \right)_0^{200},
\end{align}\vspace{-3em}

\noindent \new{where, as above, $\tau(T)$ denotes the end of growing season $\tau$ and $\tau(0) + 1$ denotes the beginning of growing season $\tau+1$.}
\new{The subscript (0) and superscript (200) indicates a lognormal distribution truncated at those values.
We used the function \texttt{urlnorm} from the Runuran package (CITATION) to generate values from the truncated lognormal distribution.}
Model parameters and notation are described in Table 1.
